---
title: "Econ Pset 3"
author: "Alex Ding"
date: '2022-05-06'
output: pdf_document
---

```{r}
# Load the ak91.csv data
df = read.csv("C:\\Users\\alexm\\Documents\\ak91.csv")
n = nrow(df)
n

# Store years of education and the weekly wage in separate variables
yrs_educ = df$YRS_EDUC
wkly_wage = df$WKLY_WAGE

```

6a:
```{r}

#Compute OLS estimates

beta = cov(wkly_wage, yrs_educ) / var(yrs_educ)
alpha = mean(wkly_wage) - mean(yrs_educ) * beta

```

6b:

```{r}
alpha
beta
```

$\hat \alpha_{n}$ = 61.1954

$\hat \beta_{n}$ = 29.6224

A brief economic interpretation of $\hat \beta_{n}$ is that an additional year of education is associated with an approximate expected change of weekly wage by 29.6224 dollars.

6c.
```{r}
# Compute BLP estimates 
blp_yx = alpha + 16 * beta
blp_yx
```
My estimate for BLP(Y $\mid$ X = 16) is 535.1538. This result differs from what I got from Problem 5b of PSet 2. This is because BLP(Y $\mid$ X = 16) does not equal E[Y $\mid$ X = 16).

6d.
```{r}
epsilon = wkly_wage - blp_yx
se_numer = sqrt(mean(epsilon^2 * (wkly_wage-mean(wkly_wage))^2))
se = (se_numer / var(wkly_wage)) / sqrt(n)
se
```
The se ($\hat \beta_{n}$) = 0.07308

6e.
The test statistic is:
```{r}
tstat = abs((beta - 31)) / se
tstat
```
The tstat is calculated to be 18.85.

6f:
The p-value is calculated by:

2 * (1- $\phi$(6.0194)) which is less than 0.00001.

p-value is less than 0.00001.

6g:
Brief economic interpretation:

On a 10% significance level, there is sufficient evidence to reject that the approximate expected change in weekly wage associated with a additional year in education is 31. This is because our p-value of ($<$ 0.0001) is less than 0.1.

7a:
```{r}
my_simplecoef <- function(y,x) {
# Compute and return estimates for alpha and beta

beta = cov(y, x) / var(x)
alpha = mean(y) - mean(x) * beta
estimates = c(alpha,beta)
return(estimates)
}
# Test the function using your solution to Problem 6
coef <- my_simplecoef(wkly_wage,yrs_educ)
coef

```

7b
```{r}
# Define a custom function to compute the blp estimates
my_simpleblp <- function (coef,x) {
blp_yx = alpha + x * beta
return(blp_yx)
}
# Test the function
mean(wkly_wage) - mean(my_simpleblp(coef, yrs_educ)) 
```

7c:
```{r}
# Define a custom function to compute the standard error
my_simplese <- function (coef,y,x) {
epsilon = y - my_simpleblp(coef, x)
se_numer = sqrt(mean(epsilon^2 * (y-mean(y))^2))
se = (se_numer / var(y)) / sqrt(n)
return(se)
}
# Test the function using your solution to Problem 6
se <- my_simplese(coef, wkly_wage, yrs_educ)
se
```
7d:
```{r}
# Define a custom function to compute the test stat and p-value
my_simpleteststat <- function (beta, se) {
  tstat = abs(beta) / se
  pval = 1 - pnorm(abs(tstat))
  tests = c(tstat,pval)
  return(tests)
}
# Test the function using your solution to Problem 6
my_simpleteststat(coef[2] - 31 , se)
```
7e.
```{r}
# Define a custom function to compute and characterize ols estimates
my_simpleols <-function (y,x) {
  my_simplecoef(y,x)
  beta = coef[2]
  se_beta = my_simplese(coef, y, x)
  tests_beta = my_simpleteststat(coef[2] - 31 , se_beta)
  ols_estimate = c(beta,se_beta,tests_beta)
  return(ols_estimate)
}
# Test the function using your solution to Problem 6
my_simpleols(wkly_wage,yrs_educ)
```

